PERFORMANCE ANALYSIS PROBLEMS IN TSN NETWORKS

### 2020-06-17 draft.tex
incorporated feedback improved figure for KNN

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Time Sensitive Networking (TSN) is a collection of standards aimed at improving the real time peformance of ethernet interface. The TSN standards, however, lack the formal latency guarantees. For that different models are proposed employing methodologies such as Network Calculus, Compositional Performance Analysis and Machine Learning. In this paper we review different performance analysis techniques at our disposal and draw comparison with respect to their methodology, completeness, accuracy and scalability.
\end{abstract}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart {T}{h}e latency requirement for industrial applications are of the order of a few millisecond, traditional networks, however, have latency of tens of millisecond \cite{LATENCY}.  Time Sensitive Networking (TSN) task group \cite{TSN} is aimed at standardizing Ultra-Low Latency (ULL) network stack on the standard ethernet interface, in order to incorporate real-time awarness on to an otherwise time agnostic ethernet data link layer protcols. There is a need of comprehensive timing analysis of such protcols to determine schedulability of dynamic time-sensitive traffic in a TSN.  There are several approaches that have been explored in order to provide the scheduling entity with the knowledge on the schedulability of a certain traffic. These methods can be broadly categorized in to two categories, mathematical based approach and simulation based approach. Mathematical approaches formulate a time aware mathematical model of a TSN protocol in order to compute latency and other characterstics such as maximum buffer size of scheduled traffic given the network traffic scenario and topology. A simulation model mimics the behaviour of the real system through a set of rules. Simulation based approach has several drawbacks such as it does not usually show the worst-case latencies and design space exploration is cumbersome.

\begin{figure}
\centering
\includegraphics[width=3.5in]{TSNFlowControl}
\caption{Flow control at an output port employing TSN CBS and TAS with and without preemption}
\label{TSN_FlowControl}
\end{figure}
A unicast or multicast connection from one end station (talker) to another end station (listener) is called as a \emph {flow}, or a \emph {stream}. TSN has defined Time Aware Shaper (TAS) \cite{IEEE802.1Qbv} and Credit Based Shaper (CBS) \cite{IEEE802.1Qav} as means to implement QoS strategies such as static priority scheduling and traffic shaping. Fig \ref{TSN_FlowControl} shows an example of traffic scheduling at an output port. For details, consult \cite{ULL}.

In section II we review two mathematical models primarily for Audio Video Bridging (AVB) traffic in a TSN network. Furthermore, we also review a Machine Learning (ML) based approach to determine schedulability of AVB traffic. In section III we carry out a discussion on the characterstics of the three models in an effort to make a qualitative comparison amongst them.

@@

\section {Analysis}
As discussed, TSN employs broad range of standards aimed at different use-cases with stringent latency requirements. In the presence of multitude of TSN standards the analysis to determine the schedulability becomes involved. In this section, we review three different techniques aimed at the timing analysis of dynamic traffic of a network.

\subsection {ML Based Analysis}
The methodology presented in \cite{ML} significantly differs from the other techniques discussed here, it does not explore underlying timing analysis to determine schedulability of a new configuration rather it makes decision based on its similarity with precomputed configurations.

\subsubsection{Model}
In order to evaluate feasibility of a new application, it is evaluated against four possible TSN configurations (or solutions). The Four TSN configurations are FIFO, Manual, Concise Priority (CP8) and Manual with traffic shaping, for details see \cite{ML}. The configurations are compatible with the standard static priority scheduling algorithms in order to compare the ML model results with so called “approximate analysis” and “precise analysis'' based off  \cite{8,9,37}.
\begin{figure}
\centering
\includegraphics[width=3.5in]{KNN}
\caption{A toy KNN model for two features, the new configuration will be characterized as `feasible' for $K<6$}
\label{KNN}
\end{figure}

A generic topology of network is given for which three classes of traffic is scheduled namely Audio, Video and Command and Control (C\&C). The ML model that is used to determine feasibility of a configuration is K Nearest Neighbors (KNN). Fig. \ref{KNN} shows a KNN model for two features X and Y. In order to classify a new configuration (marked as `?') an L2 norm is taken to determine dominant label of nearest neighbors as shown by the circles for the case of K=1 and 2.

In the ML model developed in \cite{ML}, following five features (D=5) of a configuration are selected based on the empirical results: the number of critical flows, the number of audio flows, the number of video flows, the maximum load of the network and Gini Index. Gini Index \cite{GINI} is a quantitative characterization of unbalancedness of link loads. The decision sought from the model is binary i-e if a given point in a D dimensional space connotes to either a feasible or an infeasible configuration.
\subsubsection{Results}
The training data for each solution is generated by labelling randomly generated data of the given topology by the results of `precise analysis' \cite{8,9,37} which are considered to be 100\% accurate. A training data set of N=5200 was deemed optimum.
The overall accuracy of the model reduces with increasingly complex solution. For the case of FIFO scheduling the model is a bit biased towards infeasible predictions. The reason behind that is most of the training data set contained not feasible solutions.
The accuracy of the model is worst than `approximate' analysis on average.
%The figure 6 in \cite{ML} shows the comparison of the performance of ML model against the “approximate” and “accurate” analysis in the case of Manual Scheduling, the accuracy of the model is slightly worse than the two other two analysis. However, there is a significant difference of computation time only for the case 10\^ 6 TSN configurations.
%The table 3 of \cite{ML} summarizes the accuracy result and make a comparison against the other two analyses. The interesting point to note is that the analyses do not yield False Positive Rate as opposed to the ML model, which might indicate potential room of improvement in the Approx. analysis as it is pessimistic. For the case of CP8 there is an improvement of overall accuracy in the case of ML model
\subsection {Network Calculus Based Analysis}
Network calculus (NC) is a toolset that is used in deterministic networking to compute bound on queuing, delays, buffers etc. In this paper \cite{NC} Network Calculus is used to derive an upper bound on the latency of AVB flow of TSN, which can then be used to ascertain shcedulability of the traffic.
%\subsubsection {Network Calculus Background}
%\begin{figure}
%\centering
%\includegraphics[width=3.5in]{NC_Background}
%\caption{Arrival and Departure Functions at an arbitrary node}
%\label{NC_Background}
%\end{figure}
%In network calculus, a system is characterized by two important functions that are arrival function $R(t)$ and departure function $R^*(t)$. The arrival function $R(t)$ is bounded by an arrival curve $\alpha$, which essentially is a limit on the incoming flow of the data. Similarly departure function $R^*(t)$ is bounded by a service curve $\beta$. The backlog $B(t)$ is simply a subtraction of arrival function $R(t)$ and departure function$ R^*(t)$  at a given point in time. The latency is bounded by maximum horizontal deviation between the graphs of arrival function and departure function.

%The paper primarily derives expressions for arrival and service curve in order to derive a bound on arrival and departure function which then can be used to compute latency of a particular 
%flow.
\subsubsection {Model}
The worst case analysis for an AVB traffic flow in presence of Time Triggered (TT) traffic is carried out by calculating the arrival and service curves of the flow. Arrival and service curves are functions that bound arrival and departure process respectively, for details \cite{NCBackground}. As TT traffic takes priority over any AVB flow, the paper ingeniously makes use of the arrival curves of TT traffic in order to compute service curve of the AVB flow. This makes sense because when a node is not engaged in TT traffic and credit of the particular flow is positive only then the traffic of that flow is scheduled at an output port.
\begin{figure}
\centering
\includegraphics[width=3.5in]{NC_TT}
\caption{An example of scheduled TT traffic at an output port h }
\label{NC_TT}
\end{figure}
Since TT traffic is scheduled offline assuming in a periodic fashion as show in figure \ref{NC_TT}, this period is referred to as hyperperiod, the arrival curve for TT traffic of a particular port $h$ would simply be the summation of the bits transferred in each TT traffic window in a hyperperiod.
%\begin{equation}
%\label{ttalpha}
%\alpha^h_{TT,i}(i)=\sum_{j=i}^{i+N^h-1} L_j^h.C.\Bigg \lceil \frac{t-o_{j,i}^h}{\rho_{GCL}^h}\Bigg \rceil
%\end{equation}
%The equation represents an aggregate of arrival curves where each arrival curve is with reference to each TT traffic window.

In the case of absence of IEEE  802.1Qbu, i-e no preemption of the low priority frames, the guard period needs to be catered to in the arrival curve. For the worst case the length of guard band would be either the maximum transmission time of AVB frames or just idle period if that is shorter than the aforementioned time period. Each guard band is appended in front of each TT traffic window. Following the same logic as before, the guards bands can be thought of as an extension of TT windows, since no traffic is scheduled within that time frame.
%\begingroup\makeatletter\def\f@size{8}\check@mathfonts
%$$\alpha^h_{GB + TT,i}(i)=\sum_{j=i}^{i+N^h-1} (L_j^h+L_{GB,j}^h).C.\Bigg \lceil \frac{t-o_{j,i}^h+L_{GB,j}^h-L_{GB,i}^h}{\rho_{GCL}^h}\Bigg \rceil$$\endgroup

Now, considering an AVB flow of class $M$, an interval $\Delta t$ at an output port $h$ can be broken down as: $\Delta t =\Delta t^+ + \Delta t^- + \Delta t^0$ where $\Delta t^+$ refers to the time spent in waiting, $\Delta t^-$ is the time duration of the transmission and $\Delta t^0$ refers to the time duration of TT and GB windows. As service curve is a bound on the departure function, it can be well characterized by bits served in $\Delta t^-$. This knowledge along with arrival curves of the TT and GB windows help derive the service curve for the AVB flow of class M.

The arrival curve of the AVB flow is just a summation of the bursts arrived at a node at a given point in time and their long-term rate of arrival.
%\label{avbalpha}
%\alpha^h_{AVB\_M}(t)=\sum_{\uptau_{AVB\_M_k\in h}}\sigma_{AVB\_M_k}^h+ \sum_{\uptau_{AVB\_M_k\in h}} \rho_{AVB\_M_k}^h.t
%\end{equation}
The maximum horizontal deviation of arrival and service curve of an output port $h$ draws an upper bound on the latency. The arrival curve of subsequent nodes can be estimated by summing the arrival curve of the current node with upper bound of latency for that node.
\subsubsection{Results}
The results derived for AVB analysis are shown to be scalable in presence of TT flows which are increased from 15 to 100. The analysis works for preemption where the computed Worst Case Delays (WCDs) are slightly better. The analysis also noted that the WCD of AVB flow reduce with an increased value of idle slope, naturally. The analysis also explore the effect of the routing path in a TSN topology and discovers significant impact of routing on the WCD of AVB flows.

\subsection {Compositional Performance Analysis (CPA) of AVB Ethernet Traffic}
This technique \cite{CPA} differs from the NC one primarily because of the mathematical approach towards tackling the timing analysis of the network traffic. This paper makes use of CPA toolset to derive an upper bound on the latency of a traffic stream of class AVB.
\subsubsection{CPA Background}
CPA model is composed of a set of tasks that are then processed by the a set of resources. For example an output port is a resource that executes tranmission of different flows in a network that are reqarded as tasks. Each task, $\uptau_i$,  is activated by their respective events, $q$, which are frames of a particular stream that keep the task busy referred to as busy time $B(q)$, part of the busy time is the execution time of the task on the resource associated with the task. Each task has a profile of events referred to as maximum and mimimum arrival curves, these arrival curves provide the information regarding the arrival of the events associated with the task in a given time interval.
\subsubsection{CPA based Analysis}
Busy time activated by q event is given by the equation:

@@

where: $t_{transfer}(q) = q.C_i^+$ is maximum tranfer time for back to back q events and $C$ is the time it takes for a frame to transmit at an output port. $I_{LPB}=\max\limits_{j\in lp(i)} C_j^+$ is worst-case time taken by a lower priority task that just started execution before arrival of the q-event. $I_{SPB}(a_i^q)=\sum\limits_{j\in sp(i)} (\upeta_j^+(a_i^q).C_j^+)$ is worst-case time taken by same priority events of task $\uptau_j$ arrived until arrival time of q-event of $\uptau_i$ given by $a_i^q$. $I_{HPB}(B_i^+(q,a_i^q))=\sum\limits_{j\in hp(i)} (\eta_j^+(B_i^+(q,a_i^q)-C_i^+).C_j^+)$ is worst-case time taken by a high priority task, $\uptau_j$, following the same logic as same priority task however the subtraction of $C_i^+$ is due to the non-premption nature of onging low priority task, $I_{TSB}$ is worst-case time taken by traffic shaper or insufficient credit which is dependent on other terms of the expression as blocking of $\uptau_i$ due to any reason would result in an increase of the credit. For details consult \cite{CPA}
The upper bound on latency, $l_{p}^+$, of stream $p$ is given by the summation of the worst-case response time, $R_i^+$, at each node along the path. In order to compute worst-case response time, $R_i^+$, at each node one has to find the worst-case time interval between the foremost q event arrival and the last q event departure of the same busy time characterized by $R_i^+=\max\limits_{q\in Q_i}\big\{ \max\limits_{a_i^q\in A_i}\{B_i^+(q,a_i^q)-a_i^q\}\big\}$. 
So far in the analysis, the effect of traffic shaper has not been incorporated into the calculation of arrival times of the events. For that, an upper bound on the traffic shaped by a shaper on a resource is computed which is then incorporated in to the output event model for each task. For detailed derivation consult \cite{CPA}.
\subsubsection {Results}
A general trend of much higer latencies in the case of AVB are observed which is explainable due to the maximum transmission bound imposed by shaper credit. It was seen that setting the idle slope of AVB traffic 31 times than the required bandwidth achieves the same latency as of strict priority. It was noted that in the case of linear topology the rate of increase of latencies with respect to the number of nodes was much higher than that observed in star topology. This is due to the intermediate shapers at each node that accumulated the effect of shaper at the output node.

\section{Discussion and Conclusion}
The ML paper \cite{ML} was premised around infeasibility of computational time for the case of NC based methods however the choice of ML model could have been improved as improvement in the result is only significant for the configurations $\geq 10^5$. Furthermore, it does not provide a worst-case estimate on the latencies which remains a major draw back even for a relatively accurate model. However, for a large number of configurations it provides better computation time when compared with both the NC based analyses.

The NC based analysis \cite{NC} seems to be the most comprehensive analysis amongst the reviewed techniques. The analysis also claim to provide optimistic results when compared with \cite{NCCOMPARE}, which is definitely a plus for an NC based technique. However, the analysis does not make any reference to the degree of error in the estimate. An estimate on the computational resources and time would also help in determining the applicability of the model.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Time Sensitive Networking (TSN) is a collection of standards aimed at improving the real time performance of Ethernet interface. The TSN standards, however, lack the formal latency guarantees. For that different models are proposed employing methodologies such as Network Calculus, Compositional Performance Analysis and Machine Learning. In this paper we review different performance analysis techniques at our disposal and draw comparison with respect to their methodology, completeness, accuracy and scalability.
\end{abstract}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart {T}{}he latency requirement for industrial applications are of the order of a few millisecond, traditional networks, however, have latency of tens of millisecond \cite{LATENCY}.  Time sensitive networking (TSN) task group \cite{TSN} is aimed at standardizing ultra-low latency (ULL) network stack on the standard Ethernet interface, in order to incorporate real-time awareness on to an otherwise time agnostic Ethernet data link layer protocols. There is a need of comprehensive timing analysis of such protocols to determine schedulability of dynamic time-sensitive traffic in a TSN.  There are several approaches that have been explored in order to provide the scheduling entity with the knowledge on the schedulability of a certain traffic. These methods can be broadly categorized in to two categories, mathematical based approach and simulation based approach. Mathematical approaches formulate a time aware mathematical model of a TSN protocol in order to compute latency and other characteristics such as maximum buffer size of scheduled traffic given the network traffic scenario and topology. A simulation model mimics the behavior of the real system through a set of rules. Simulation based approach has several drawbacks such as it does not usually show the worst-case latencies and design space exploration is cumbersome.

\begin{figure}
\centering
\includegraphics[width=3.5in]{TSNFlowControl}
\caption{Flow control at an output port employing TSN CBS and TAS with and without preemption}
\label{TSN_FlowControl}
\end{figure}
A unicast or multicast connection from one end station (talker) to another end station (listener) is called as a \emph {flow}, or a \emph {stream}. TSN has defined time aware shaper (TAS) \cite{IEEE802.1Qbv} and credit based shaper (CBS) \cite{IEEE802.1Qav} as means to implement QoS strategies such as static deterministic scheduling and traffic shaping. Fig \ref{TSN_FlowControl} shows an example of traffic scheduling at an output port. For more details of standards in TSN, consult \cite{ULL}.

In section II we review a machine learning (ML) based approach to determine schedulability of AVB traffic. Furthermore, we review two mathematical models primarily for audio video bridging (AVB) traffic in a TSN network. In section III we carry out a discussion on the characteristics of the three models in an effort to make a qualitative comparison among them.

@@

\section {Analysis}
TSN employs broad range of standards aimed at different use-cases with stringent latency requirements. In the presence of multitude of TSN standards the analysis to determine the schedulability becomes involved. In this section, we review three different techniques aimed at the timing analysis of dynamic traffic of a network.

\subsection {ML Based Analysis}
The methodology presented in \cite{ML} significantly differs from the other techniques discussed here, it does not carry out timing analysis of a new configuration to determine schedulability rather it makes decision based on the similarity of the configuration with pre-computed configurations.

\subsubsection{Model}
In order to evaluate feasibility of a new application (configuration), it is evaluated against four possible TSN configurations (or solutions). The four TSN configurations are FIFO, manual, concise priority (CP8) and manual with traffic shaping, for details see \cite{ML}. The configurations are compatible with the standard static priority scheduling algorithms in order to compare the ML model results with so called “approximate analysis” and “precise analysis'' based off  \cite{8,9,37}.
\begin{figure}
\centering
\includegraphics[width=3.5in]{KNN}
\caption{A toy KNN model for two features, the new configuration will be characterized as `feasible' for $K<6$}
\label{KNN}
\end{figure}

A generic topology of a network is given for which three classes of network traffic is scheduled namely audio, video and command and control (C\&C). The ML model that is used to determine feasibility of a configuration is K nearest neighbors (KNN). Fig. \ref{KNN} shows a KNN model for two features X and Y. In order to classify a new configuration (marked as `?') an L2 norm is taken to determine dominant label of nearest neighbors as shown by the circles for the case of K=1 and 2.

In the ML model developed in \cite{ML}, following five features (D=5) of a configuration are chosen based on empirical results: the number of critical flows, the number of audio flows, the number of video flows, the maximum load of the network and Gini index. Gini Index \cite{GINI} is a quantitative characterization of unbalancedness of link loads. The decision sought from the model is binary i-e if a given point in a D dimensional space connotes to either a feasible or an infeasible configuration.
\subsubsection{Results}
The training data (for instance shapes in figure \ref{KNN}) for each solution is generated by labeling (for instance color of the shapes) randomly generated data of the given topology by the results of `precise analysis' \cite{8,9,37} which are considered to be 100\% accurate. A training data set of N=5200 was deemed optimum.
The overall accuracy of the model reduces with increasingly complex solution. For the case of FIFO scheduling the model is a bit biased towards infeasible predictions. The reason behind that is most of the training data set contained not feasible solutions, the bias disappears for the other solutions as coverage for feasible feature space increases. The accuracy of the model is worst than `approximate' analysis on average, considering `precise analysis' as an 100\% accurate reference.
%The figure 6 in \cite{ML} shows the comparison of the performance of ML model against the “approximate” and “accurate” analysis in the case of Manual Scheduling, the accuracy of the model is slightly worse than the two other two analysis. However, there is a significant difference of computation time only for the case 10\^ 6 TSN configurations.
%The table 3 of \cite{ML} summarizes the accuracy result and make a comparison against the other two analyses. The interesting point to note is that the analyses do not yield False Positive Rate as opposed to the ML model, which might indicate potential room of improvement in the Approx. analysis as it is pessimistic. For the case of CP8 there is an improvement of overall accuracy in the case of ML model
\subsection {Network Calculus Based Analysis}
Network calculus (NC) is a theory that is used in deterministic networking to compute bound on queuing, delays, buffers etc. In\cite{NC}, NC is used to derive an upper bound on the latency of AVB flows of TSN, which can then be used to ascertain schedulability of the traffic.
%\subsubsection {Network Calculus Background}
%\begin{figure}
%\centering
%\includegraphics[width=3.5in]{NC_Background}
%\caption{Arrival and Departure Functions at an arbitrary node}
%\label{NC_Background}
%\end{figure}
%In network calculus, a system is characterized by two important functions that are arrival function $R(t)$ and departure function $R^*(t)$. The arrival function $R(t)$ is bounded by an arrival curve $\alpha$, which essentially is a limit on the incoming flow of the data. Similarly departure function $R^*(t)$ is bounded by a service curve $\beta$. The backlog $B(t)$ is simply a subtraction of arrival function $R(t)$ and departure function$ R^*(t)$  at a given point in time. The latency is bounded by maximum horizontal deviation between the graphs of arrival function and departure function.

%The paper primarily derives expressions for arrival and service curve in order to derive a bound on arrival and departure function which then can be used to compute latency of a particular 
%flow.
\subsubsection {Model}
The worst case analysis for an AVB traffic flow in presence of time triggered (TT) traffic is carried out by calculating the arrival and service curves of the flow. Arrival and service curves are functions that bound arrival and departure process respectively \cite{NCBackground}. As TT traffic takes priority over any AVB flow, the paper ingeniously makes use of the arrival curves of TT traffic in order to compute service curve of the AVB flow. This makes sense because when a node is not engaged in TT traffic and credit of the particular flow is positive only then the AVB traffic is scheduled at an output port.
\begin{figure}
\centering
\includegraphics[width=3.5in]{NC_TT}
\caption{An example of scheduled TT traffic at an output port h }
\label{NC_TT}
\end{figure}
Since TT traffic is scheduled offline assuming in a periodic fashion as show in figure \ref{NC_TT}, this period is referred to as hyperperiod, the arrival curve for TT traffic of a particular port $h$ would simply be the summation of the bits transferred in each TT traffic window in a hyperperiod.
%\begin{equation}
%\label{ttalpha}
%\alpha^h_{TT,i}(i)=\sum_{j=i}^{i+N^h-1} L_j^h.C.\Bigg \lceil \frac{t-o_{j,i}^h}{\rho_{GCL}^h}\Bigg \rceil
%\end{equation}
%The equation represents an aggregate of arrival curves where each arrival curve is with reference to each TT traffic window.

In the case of absence of IEEE  802.1Qbu, i-e no preemption of the low priority frames, the guard period needs to be catered to in the arrival curve. For the worst case the length of guard band (GB) would be either the maximum transmission time of AVB frames or just idle period if that is shorter than the aforementioned time period. Each GB is appended in front of each TT traffic window. Following the same logic as before, the GBs can be thought of as an extension of TT windows, since no traffic is scheduled within that time frame.
%\begingroup\makeatletter\def\f@size{8}\check@mathfonts
%$$\alpha^h_{GB + TT,i}(i)=\sum_{j=i}^{i+N^h-1} (L_j^h+L_{GB,j}^h).C.\Bigg \lceil \frac{t-o_{j,i}^h+L_{GB,j}^h-L_{GB,i}^h}{\rho_{GCL}^h}\Bigg \rceil$$\endgroup

Now, considering the service curve of AVB traffic, an interval $\Delta t$ at an output port $h$ can be broken down as: $\Delta t =\Delta t^+ + \Delta t^- + \Delta t^0$ where $\Delta t^+$ refers to the time spent in waiting, $\Delta t^-$ is the time duration of the transmission and $\Delta t^0$ refers to the time duration of TT and GB windows. As service curve is a bound on the departure function, it can be well characterized by bits served in $\Delta t^-$. This knowledge along with arrival curves of the TT and GB windows help derive the service curve for the AVB flow.

The arrival curve of the AVB traffic is just a summation of the bursts arrived at a node at a given point in time and their long-term rate of arrival.
%\label{avbalpha}
%\alpha^h_{AVB\_M}(t)=\sum_{\uptau_{AVB\_M_k\in h}}\sigma_{AVB\_M_k}^h+ \sum_{\uptau_{AVB\_M_k\in h}} \rho_{AVB\_M_k}^h.t
%\end{equation}
The maximum horizontal deviation of arrival and service curve of AVB traffic, at a output port $h$, draws an upper bound on the latency.
\subsubsection{Results}
The results derived for AVB analysis are shown to be scalable in the presence of varying number of TT flows. The analysis compared the worst case delay (WCD) of preemption mode with that of non-preemption mode. The analysis also noted that the WCD of AVB flow reduced with an increased value of the idle slope, naturally. The analysis also explored the effect of the routing path in a TSN topology and discovered significant impact of routing on the WCD of AVB flows.

\subsection {Compositional Performance Analysis (CPA) of AVB Ethernet Traffic}
Compositional performance analysis (CPA) \cite{CPA} is another mathematical approach different from NC because of the mathematical approach towards tackling the timing analysis of network traffic. The paper \cite{CPA} makes use of CPA theory to derive an upper bound on the latency of a traffic flow of class AVB.
\subsubsection{CPA Background}
CPA model is composed of a set of tasks that are then processed by a set of resources. For example an output port is a resource that executes transmission of different flows in a network that are regarded as tasks. Each task, $\uptau_i$,  is activated by their respective events, $q$, which are frames of a particular stream that keep the task busy referred to as busy time $B(q)$, part of the busy time is attributed to scheduling of the task on the resource associated with the task, while the rest is attributed to waiting times of different types. Each task has a profile of events referred to as maximum and minimum arrival curves, these arrival curves provide the information regarding the arrival of the events associated with the task in a given time interval.
\subsubsection{CPA based Analysis}
Busy time activated by a q event is given by the equation:

@@

where: $t_{transfer}(q) = q.C_i^+$ is maximum transfer time for back to back q events and $C$ is the time it takes for a frame to transmit at an output port. $I_{LPB}=\max\limits_{j\in lp(i)} C_j^+$ is worst-case time taken by a lower priority task that just started execution before arrival of the q-event. $I_{SPB}(a_i^q)=\sum\limits_{j\in sp(i)} (\upeta_j^+(a_i^q).C_j^+)$ is worst-case time taken by same priority events of task $\uptau_j$ arrived until arrival time of q-event of $\uptau_i$ given by $a_i^q$. $I_{HPB}(B_i^+(q,a_i^q))=\sum\limits_{j\in hp(i)} (\eta_j^+(B_i^+(q,a_i^q)-C_i^+).C_j^+)$ is worst-case time taken by a high priority task, $\uptau_j$, following the same logic as same priority task however the subtraction of $C_i^+$ is due to the non-preemption nature of ongoing low priority task, $I_{TSB}$ is worst-case time taken by traffic shaper or insufficient credit which is dependent on other terms of the expression as blocking of $\uptau_i$ due to any reason would result in an increase of the credit. For details consult \cite{CPA}
The upper bound on latency, $l_{p}^+$, of stream $p$ is given by the summation of the worst-case response time, $R_i^+$, at each node along the path. In order to compute worst-case response time, $R_i^+$, at each node one has to find the worst-case time interval between the foremost q event arrival and the last q event departure of the same busy time characterized by $R_i^+=\max\limits_{q\in Q_i}\big\{ \max\limits_{a_i^q\in A_i}\{B_i^+(q,a_i^q)-a_i^q\}\big\}$. 
So far in the analysis, the effect of traffic shaper has not been incorporated into the calculation of arrival times of the events. For that, an upper bound on the traffic shaped by a shaper on a resource is computed which is then incorporated in to the output event model for each task. For detailed derivation consult \cite{CPA}.
\subsubsection {Results}
A general trend of much higher latencies in the case of AVB flow, when compared with strict priority flow, are observed which is explainable due to the maximum transmission bound imposed by shaper credit. It was seen that setting the idle slope of AVB traffic 31 times than the required bandwidth achieves the same latency as of strict priority. It was noted that in the case of linear topology the rate of increase of latencies with respect to the number of nodes was much higher than that observed in star topology. This is due to the intermediate shapers at each node that accumulated the effect of shaper at the output node.

\section{Discussion and Conclusion}
The ML paper \cite{ML} was premised around infeasibility of computational time for the case of NC based methods however the choice of ML model could have been improved as improvement in the result is only significant for TSN configurations $\geq 10^5$. Furthermore, it does not provide a worst-case estimate on the latencies which remains a major draw back even for a relatively accurate model. However, for a large number of configurations it provides better computation time when compared with both the mathematical analyses.

The NC based analysis \cite{NC} seems to be the most comprehensive analysis among the reviewed techniques. The analysis also claim to provide optimistic results when compared with \cite{NCCOMPARE}, which is definitely a plus for an NC based technique. However, the analysis does not make any reference to the degree of pessimism in the estimate. An estimate on the computational resources and time would also help in determining the applicability of the model.
